NLIZE: A Perturbation-Driven Visual Interrogation Tool for Analyzing and Interpreting Natural Language Inference Models: Supplemental Materials

### 2018-06-22 paper/attention.tex
update paper

Even though the attention does not explicitly encode any grammar, it often highlights essential words in the sentence structure.
%
To help the researcher better understand the relationship between attention and sentence structure, as illustrated in Fig.~\ref{fig:attentionVis}($a_{1}$), we overlay the grammar dependency tree~\cite{Nivre2005} structure next to the sentence.
%
Since the dependency tree encodes the word importance information in a hierarchical manner, it is very suitable for sentence simplification tasks.
Therefore, we utilize the grammar dependency tree to simplify the attention representations to combat the visual clutter when examining long sentences (see Fig.~\ref{fig:attentionVis}(d)). A simplification example is shown in  Fig.~\ref{fig:attentionVis}(d)(e)(f).

>>>>

Even though the attention does not explicitly encode any grammar, it often highlights essential words in the sentence structure.
%
To help the researcher better understand the relationship between attention and sentence structure, as illustrated in Fig.~\ref{fig:attentionVis}($a_{1}$), we overlay the grammar dependency tree~\cite{Nivre2005} structure next to the sentence.
%
Since the dependency tree encodes the word importance information in a hierarchical manner, it is very suitable for sentence simplification tasks.
Here, we utilize the grammar dependency tree to trim the decorative structure to shorten the sentence to combat the visual clutter when examining long sentences (see Fig.~\ref{fig:attentionVis}(d)). A simplification example is shown in  Fig.~\ref{fig:attentionVis}(d)(e)(f).

### 2018-06-22 paper/intro.tex
update paper

One recent advance in neural natural language process models is the
introduction of \emph{attention
  mechanisms}~\cite{bahdanau2014neural,VaswaniShazeerParmar2017}
(Section~\ref{sec:attention}). Intuitively, attention asks which parts of the
input are deemed more important for making a prediction. The attention is often represented via weights for individual words or pairs of words (i.e., the alignment between words in different sentences).
%
%\shusen{seems repeated some of the previous points}
There have been many speculations on how attention works in various models.
Using the proposed tool, the domain experts can study how changes in sentence input, attention, or prediction affect each other, which helps them develop deeper intuitive and alternative hypotheses.
%we can perturb the attention and observer the change in the prediction, perturb the sentence input and observe the effects on the attention. And more importantly, we also allow the user to change prediction, where we apply a constrained optimization to find the least amount of parameter update to produce the user assign label.
%Another word,
%Such type of exploration is not only essential to interpret the language inference problem, but can also be generalized to other NLP model utilized the attention mechanisms.
Moreover, we propose to enhance the standard visual encoding (e.g., as a bipartite graph or as a matrix) of the attention matrix by overlaying sentence linguistic structure to allow grammar-guided simplification of the visual representation.
%
Finally, as discussed in Section~\ref{sec:grammarAttention}, the ability to examine how attention corresponds to the grammatical structure also enables domain experts to speculate about the potential benefits of including the linguistic structure in the design of the attention component of the model.

>>>>

One recent advance in neural natural language process models is the
introduction of \emph{attention
  mechanisms}~\cite{bahdanau2014neural,VaswaniShazeerParmar2017}
(Section~\ref{sec:attention}). Intuitively, attention asks which parts of the
input are deemed more important for making a prediction. The attention is often represented via weights for individual words or pairs of words (i.e., the alignment between words in different sentences).
%
%\shusen{seems repeated some of the previous points}
There have been many speculations on how attention works in various models.
The proposed tool introduces a perturbation-driven visual analytics enviroment, where the domain experts can study how changes in sentence input, attention, or prediction affect each other, which helps them develop deeper intuitive and alternative hypotheses.
%we can perturb the attention and observer the change in the prediction, perturb the sentence input and observe the effects on the attention. And more importantly, we also allow the user to change prediction, where we apply a constrained optimization to find the least amount of parameter update to produce the user assign label.
%Another word,
%Such type of exploration is not only essential to interpret the language inference problem, but can also be generalized to other NLP model utilized the attention mechanisms.
In addition, we propose to enhance the standard visual encoding (e.g., as a bipartite graph or as a matrix) of the attention matrix by overlaying sentence linguistic structure to allow grammar-guided simplification of the visual representation.
%
Finally, as discussed in Section~\ref{sec:grammarAttention}, the ability to examine how attention corresponds to the grammatical structure also enables domain experts to speculate about the potential benefits of including the linguistic structure in the design of the attention component of the model.

### 2018-06-22 paper/pipeline.tex
update paper

\begin{figure}[htbp]
\centering
\vspace{-2mm}
 \includegraphics[width=0.95\linewidth]{pipelineView}
 \vspace{-2mm}
 \caption{
The pipeline view provides the visual representation of the three-stage (encoder, attention, classifier) model. 
%
In the proposed tool, we allow model parameters to be updated to correct a prediction error. 
The optimization hyperparameters are shown in (a), and each of the stages is illustrated in (b), where the user can click on the blue rectangle marked with the word ``parameter'' to enable or disable its update (the legend is shown in (d)). In (c), we select whether we want to use the current pipeline setting as displayed or try all the pipeline configuration combinations as the parameter update operation can be enabled or disabled for each stage.
 }
\label{fig:pipelineView}
\vspace{-2mm}
\end{figure}

>>>>

\begin{figure}[htbp]
\centering
\vspace{-2mm}
 \includegraphics[width=0.95\linewidth]{pipelineView}
 \vspace{-2mm}
 \caption{
The pipeline view provides the visual representation of the three-stage (encoder, attention, classifier) of the NLI model. 
%
In the proposed tool, we allow model parameters to be updated to correct a prediction error via a constrained optimization.
The hyperparameters for this optimization are shown in (a). 
In (b), we utilizes a graphical representation for each stage of the pipeline.
% where the user can click on the blue rectangle marked with the word ``parameter'' to enable or disable its update (the legend is shown in (d)). 
In (c), the user can select whether to use the current pipeline configuration as displayed or try all the pipeline configuration combinations for the optimization. %whether we want to use the current pipeline configuration as displayed or try all the pipeline configuration combinations as the parameter update operation can be enabled or disabled for each stage.
}
\label{fig:pipelineView}
\vspace{-2mm}
\end{figure}

### 2018-06-22 paper/predSummary.tex
update paper

\begin{figure}[htbp]
\centering
\vspace{-2mm}
 \includegraphics[width=1.0\linewidth]{summaryView}
 \vspace{-6mm}
 \caption{
%Summarize the prediction results and sensitivity of the entire development set, and provide an explorative interface to select an example of interest.
We summarize all the prediction results in (a), where the green block indicates correct predictions, the orange block indicates wrong predictions, and the label (e.g., E/E, E/N) shows the ground truth label and the predicted label (N-Neural, E-Entailment, C-Contradiction).
%
The user can select the treemap node to focus on the specific type of scenarios (e.g., E/E) and reveal the histogram (b) and scatterplot (c) for displaying the selected subset.
The selection can be further narrowed down by selecting the bin in the histogram.
In (c) and (d), each point corresponds to one sentence pair.
 }
 \vspace{-2mm}
\label{fig:summaryView}
\end{figure}

@@

>>>>

@@

\begin{figure}[htbp]
\centering
\vspace{-2mm}
 \includegraphics[width=1.0\linewidth]{summaryView}
 \vspace{-6mm}
 \caption{
%Summarize the prediction results and sensitivity of the entire development set, and provide an explorative interface to select an example of interest.
We summarize all the prediction results in (a), where the green block indicates correct predictions, the orange block indicates wrong predictions, and the label (e.g., E/E, E/N) shows the ground truth label and the predicted label (N-Neural, E-Entailment, C-Contradiction).
%
The user can click on the treemap node to focus on the specific type of scenarios (e.g., E/E) to automatically reveal the histogram (b) and scatterplot (c) for displaying the selected subset.
The selection can be further narrowed down by selecting the bin in the histogram.
In (c) and (d), each point corresponds to one sentence pair.
 }
 \vspace{-2mm}
\label{fig:summaryView}
\end{figure}

### 2018-06-22 paper/prediction.tex
update response letter
finish response to review 4
remove related works
update paper

\subsection{Prediction View}
\label{sec:prediction}
For a given sentence pair, the model predicts a discrete probability distribution of the three labels (neutral, contradiction, and entailment).
%
In the prediction view, as illustrated in Fig.~\ref{fig:predictionView}(a), a prediction probability is encoded as a point in the barycentric coordinate system of the triangle.
Let $\mathbf{C_1}$,$\mathbf{C_2}$,$\mathbf{C_3}$ be the three points defining a triangle, and let $p_1$, $p_2$, $p_3$ be the probabilities the label is neutral, contradiction or entailment. The coordinate of the prediction $\mathbf{C_p}$ in the triangle is computed as $\mathbf{C_p} = p_1\mathbf{C_1}+p_2\mathbf{C_2}+p_3\mathbf{C_3}$.
In the triangle, the distinctly colored background (grey, light grey, and light green) indicates the regions that correspond to different labels. The prediction result for the original sentence pair is represented by the larger yellow circle, whereas the smaller gray circles illustrate the perturbed sentence pairs. A density contour of the prediction is computed to emphasize the highly cluttered areas and distinguish the outliers.

\begin{figure}[htbp]
\centering
\vspace{-2mm}
 \includegraphics[width=1.0\linewidth]{predictionView}
 \vspace{-5mm}
 \caption{
In the prediction view, the prediction is encoded as a point in the barycentric coordinate system of the triangle shown in (a).
 %
% The prediction result for the original sentence pair is represented by a larger yellow circle and the prediction of perturbed pairs are illustrated by smaller grey circles.
A density contour of the prediction is computed to emphasize the highly cluttered areas and distinguish the outliers.
As illustrated in (b), the predicted label does not match the ground truth. Therefore, we apply a label reassignment operation, which triggers a model update.
% when press and drag the circle (prediction), the user is presented with the three option corresponds to the labels. The dotted line and arrow indicate the newly assigned label.
 }
  \vspace{-2mm}
\label{fig:predictionView}
\end{figure}

>>>>

\subsection{Prediction View}
\label{sec:prediction}
For a given sentence pair, the model predicts a discrete probability distribution of the three labels (neutral, contradiction, and entailment).
%
In the prediction view, as illustrated in Fig.~\ref{fig:predictionView}(a), a prediction probability is encoded as a point in the barycentric coordinate system of the triangle.
Let $\mathbf{C_1}$,$\mathbf{C_2}$,$\mathbf{C_3}$ be the three points defining a triangle, and let $p_1$, $p_2$, $p_3$ be the probabilities the label is neutral, contradiction or entailment. The coordinate of the prediction $\mathbf{C_p}$ in the triangle is computed as $\mathbf{C_p} = p_1\mathbf{C_1}+p_2\mathbf{C_2}+p_3\mathbf{C_3}$.
In the triangle, the distinctly colored background (grey, light grey, and light green) indicates the regions that correspond to different labels. The prediction result for the original sentence pair is represented by the larger yellow circle, whereas the smaller gray circles illustrate the perturbed sentence pairs. A density contour of the prediction is computed to emphasize the highly cluttered areas and distinguish the outliers.
%\shusen{
The pattern of the perturbed pairs' prediction directly convey the stability of the model for the given instance. 
We should also take the length of the sentence into consideration as more nouns and verbs will likely lead to more perturbed inputs.
%}

\begin{figure}[htbp]
\centering
\vspace{-2mm}
 \includegraphics[width=1.0\linewidth]{predictionView}
 \vspace{-5mm}
 \caption{
In the prediction view, the prediction is encoded as a point in the barycentric coordinate system of the triangle shown in (a).
 %
% The prediction result for the original sentence pair is represented by a larger yellow circle and the prediction of perturbed pairs are illustrated by smaller grey circles.
A density contour of the prediction is computed via kernel density estimation to emphasize the highly cluttered areas and distinguish the outliers.
As illustrated in (b), the predicted label does not match the ground truth. Therefore, we apply a label reassignment operation, which triggers a model update.
% when press and drag the circle (prediction), the user is presented with the three option corresponds to the labels. The dotted line and arrow indicate the newly assigned label.
}
\vspace{-2mm}
\label{fig:predictionView}
\end{figure}
